Wrangle

### 2023-10-06 continuous-time.qmd
Cross reference sections (#11)
WIP

---
title: "Continuous time"
---

@@

In the main analysis in [Analyze]() we focused on contrasting the mood at the beginning of a session with the average mood during session. In this document we focus on two additional research questions:

@@

library(scales)
library(cmdstanr)
library(posterior)
library(brms)
library(patchwork)
library(tidyverse)
library(splines)
library(emmeans)
library(lme4)
library(broom.mixed)
library(knitr)
library(rstanarm)
source("R/functions.R")
source("R/fit_cached.R")
```

```{r}
#| label: options
#| cache: false

N_CORES <- as.numeric(Sys.getenv("N_CORES", unset = 1))
N_THREADS <- as.numeric(Sys.getenv("N_THREADS", unset = 2))
N_ITER <- as.numeric(Sys.getenv("N_ITER", unset = 2000))
N_SUBSET <- as.numeric(Sys.getenv("N_SUBSET"))
options(
  mc.cores = N_CORES,
  brms.backend = "cmdstanr"
)
theme_set(
  theme_linedraw(base_size = 10) +
    theme(
      strip.background = element_rect(
        fill = NA,
        color = NA
      ),
      strip.text = element_text(
        colour = "black", hjust = 0
      ),
      axis.text = element_text(
        size = rel(0.75)
      )
    )
)
scale_x_hours <- list(
  scale_x_continuous(
  breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3), 
  labels = c(0, "15m", "30m", "45m", "1h", "2h", "3h")
  )
)
```

## Data
Load the data and remove observations that occurred more than 3 hours into a session.
```{r}
#| label: load-data
dat <- read_rds("data/data.rds")

dat <- dat |>
  filter(
    !is.na(mood),
    hours <= 3
  ) |>
  mutate(
    cl = case_when(
      mood == 0 ~ "left",
      mood == 100 ~ "right",
      TRUE ~ "none"
    ),
  )
if (!is.na(N_SUBSET)) {
  pid_subset <- sample(unique(dat$pid), N_SUBSET)
  dat <- filter(dat, pid %in% pid_subset) |> 
    droplevels()
}
```

@@

```{r}
#| label: fig-plot-curve
#| fig-cap: Estimated mood during a play session
#| fig-height: 3
#| fig-width: 8
hours <- seq(0, 3, length.out = 100)
emm1 <- emmeans(
  fit,
  ~hours,
  at = list(hours = hours),
  lmer.df = "asymptotic"
)
p_curve <- as.data.frame(emm1) |>
  ggplot(aes(hours, emmean)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  labs(
    x = "Session duration (hours)",
    y = "Mood (0 - 100)"
  ) +
  scale_x_hours
p_curve
```

@@

```{r}
#| label: fig-plot-change
#| fig-cap: Estimated difference in mood over a session compared to the start of a session
#| fig-height: 3
#| fig-width: 8
d_change <- emmeans::contrast(
  emm1,
  method = "trt.vs.ctrl",
  ref = "hours0"
) |>
  confint() |>
  as.data.frame() |>
  mutate(
    hours = hours[-1]
  )

p_change <- d_change |>
  ggplot(aes(hours, estimate)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  lims(y = c(0, 10)) +
  labs(x = "Session duration (hours)") +
  scale_x_hours
p_change
```

@tbl-change-contrast shows the estimated change from baseline after 15min, 30 min, 1 hour, 2 hours, and 3 hours.
```{r}
#| label: tbl-change-contrast
#| tbl-cap: Estimated change from baseline
emmeans(
  fit,
  ~hours,
  at = list(hours = c(0, 15 / 60, 0.5, 1, 2, 3)),
  lmer.df = "asymptotic"
) |>
  emmeans::contrast(
    method = "trt.vs.ctrl",
    ref = "hours0",
  ) |>
  confint() |>
  kable(
    digits = 2
  )
```

@@

Plot the interaction at different pretest values. As shown in @fig-lmm-pre-interaction, this interaction model shows a pattern similar to @fig-intercept-sds where those who report lower mood at the beginning of a session also improve more. 
```{r}
#| label: fig-lmm-pre-interaction
#| fig-cap: Interaction between pretest mood and change, shown at -2 SD, -1, SD, the mean, and +1 SD.
#| fig-height: 3
#| fig-width: 8
emm <- emmeans(
  fit3,
  ~ hours + pre,
  at = list(
    hours = seq(0, 3, length.out = 100),
    pre = pre_values
  ),
  lmer.df = "asymptotic"
)
p_pre_x <- as.data.frame(emm) |>
  mutate(
    pre = factor(
      round(pre), 
      labels = pre_labels_x
      )
  ) |>
  ggplot(
    aes(
      hours,
      emmean,
      group = pre,
      color = pre
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      fill = pre,
      color = NULL
    ),
    alpha = 0.25,
    show.legend = FALSE
  ) +
  color_scale_x +
  scale_x_hours

@@

Let's also calculate the within-session change for the interaction model.
```{r}
#| label: fig-lmm-pre-interaction-change
#| fig-cap: Within-session change from the interaction model
#| fig-height: 2.5
#| fig-width: 8
source("R/get_change_by_pre.R")
d_x_change <- data.frame(
  pre = round(pre_values),
  label = pre_labels_x
) |>
  pmap(\(pre, label) get_change_by_pre(
    pre,
    label,
    fit3,
    hours = seq(0, 3, length.out = 100)
  )) |>
  bind_rows()

@@

### Manuscript figures
Create figures for the manuscript.
```{r}
#| label: fig-ms-interaction
#| fig-height: 4
#| fig-width: 8
#| fig-cap: Figure included in manuscript

scale_x_hours2 <- list(
  scale_x_continuous(
  breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3), 
  labels = c(0, "", "30m", "", "1h", "2h", "3h")
  )
)
p_tmp <- (
  (
    p_pre_x + 
      facet_wrap(~pre, ncol = 4) +
      scale_x_hours2 +
      labs(y = "Mood", x = NULL) +
      theme(plot.margin = margin(0, 0, 0, 0, "pt"))
    ) / 
    (
      p_pre_x_change +
        scale_x_hours2 +
        labs(y = "Change", x = "Duration") +
        theme(plot.margin = margin(0, 0, 0, 0, "pt"))
      ) 
    ) +
   plot_annotation(tag_levels = "A") +
  plot_layout(
    heights = c(2, 2),
    widths = c(8, 8),
    guides = "collect"
  ) & theme(legend.position = "bottom") 
p_tmp
```

```{r}
#| label: ggsave-fig-ms-interaction
#| echo: false
ggsave(
  "figures/fig-pre-x-change.png", 
  p_tmp, 
  width = 7, 
  height = 3.5
  )
```

@@

```{r}
#| label: fig-ms-no-interaction
#| fig-height: 4
#| fig-width: 8
#| fig-cap: Figure shown in manuscript
p_tmp <- (
  p_curve + labs(x = NULL, y = "Mood")
  ) / (
  p_change + labs(x = "Duration", y = "Change")
  ) + plot_annotation(tag_levels = "A") 
p_tmp
```

```{r}
#| label: ggsave-fig-ms-no-interaction
#| echo: false
ggsave(
  "figures/fig-change.png", 
  p_tmp, 
  width = 6, 
  height = 3
  )
```

@@

get_trend_by_model <- function(fit, model_label) {
  emm <- emmeans(
    fit,
    ~ hours + pre,
    at = list(
      hours = seq(0, 3, length.out = 100),
      pre = pre_values
    ),
    lmer.df = "asymp"
  )
  as.data.frame(emm) |>
    mutate(
      pre = factor(
        round(pre),
        labels = pre_labels_x,
      ),
      model = model_label
    )
}
tmp_p <- map2(
  tmp_models,
  names(tmp_models),
  \(fit, lab) get_trend_by_model(fit, lab)
) |>
  bind_rows()
p_pre_x_sens <- tmp_p |>
  ggplot(
    aes(
      hours,
      emmean,
      group = interaction(model, pre),
      color = model,
      linetype = pre
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      fill = model,
      color = NULL
    ),
    alpha = 0.10,
    show.legend = FALSE
  ) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  scale_x_hours
p_pre_x_sens
```

>>>>

# Continuous time {#sec-continuous-time}

@@

In the main analysis in @sec-analyze we focused on contrasting the mood at the beginning of a session with the average mood during session. In this document we focus on two additional research questions:

@@

library(scales)
library(cmdstanr)
library(posterior)
library(brms)
library(patchwork)
library(tidyverse)
library(splines)
library(emmeans)
library(lme4)
library(broom.mixed)
library(knitr)
library(rstanarm)
source("R/functions.R")
source("R/fit_cached.R")
source("R/common.R")
```

## Data
Load the data.
```{r}
#| label: load-data

dat <- read_rds("data/data.rds")

if (!is.na(N_SUBSET)) {
  pid_subset <- sample(unique(dat$pid), N_SUBSET)
  dat <- filter(dat, pid %in% pid_subset) |> 
    droplevels()
}
```

@@

```{r}
#| label: fig-plot-curve
#| fig-cap: Estimated mood during a play session
#| fig-height: 3
#| fig-width: 8

hours <- seq(0, 3, length.out = 100)
emm1 <- emmeans(
  fit,
  ~hours,
  at = list(hours = hours),
  lmer.df = "asymptotic"
)
p_curve <- as.data.frame(emm1) |>
  ggplot(aes(hours, emmean)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  labs(
    x = "Session duration (hours)",
    y = "Mood (0 - 1)"
  ) +
  scale_x_hours
p_curve
```

@@

```{r}
#| label: fig-plot-change
#| fig-cap: Estimated difference in mood over a session compared to the start of a session
#| fig-height: 3
#| fig-width: 8

d_change <- emmeans::contrast(
  emm1,
  method = "trt.vs.ctrl",
  ref = "hours0"
) |>
  confint() |>
  as.data.frame() |>
  mutate(
    hours = hours[-1]
  )

p_change <- d_change |>
  ggplot(aes(hours, estimate)) +
  scale_y_continuous(
    expand = expansion(c(0.1, 0.1))
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  labs(x = "Session duration (hours)") +
  scale_x_hours
p_change
```

@tbl-change-contrast shows the estimated change from baseline after 15min, 30 min, 1 hour, 2 hours, and 3 hours.
```{r}
#| label: tbl-change-contrast
#| tbl-cap: Estimated change from baseline
emmeans(
  fit,
  ~hours,
  at = list(hours = c(0, 15 / 60, 0.5, 1, 2, 3)),
  lmer.df = "asymptotic"
) |>
  emmeans::contrast(
    method = "trt.vs.ctrl",
    ref = "hours0",
  ) |>
  confint() |>
  kable(
    digits = 3
  )
```

@@

Plot the interaction at different pretest values. As shown in @fig-lmm-pre-interaction, this interaction model shows a pattern similar to @fig-intercept-sds where those who report lower mood at the beginning of a session also improve more. 
```{r}
#| label: fig-lmm-pre-interaction
#| fig-cap: Interaction between pretest mood and change, shown at -2 SD, -1, SD, the mean, and +1 SD.
#| fig-height: 3
#| fig-width: 8
emm <- emmeans(
  fit3,
  ~ hours + pre,
  at = list(
    hours = seq(0, 3, length.out = 100),
    pre = pre_values
  ),
  lmer.df = "asymptotic"
)
p_pre_x <- as.data.frame(emm) |>
  mutate(
    pre = factor(
      pre,
      labels = pre_labels_x
      )
  ) |>
  ggplot(
    aes(
      hours,
      emmean,
      group = pre,
      color = pre
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      fill = pre,
      color = NULL
    ),
    alpha = 0.25,
    show.legend = FALSE
  ) +
  color_scale_x +
  scale_x_hours

@@

Let's also calculate the within-session change for the interaction model.
```{r}
#| label: fig-lmm-pre-interaction-change
#| fig-cap: Within-session change from the interaction model
#| fig-height: 2.5
#| fig-width: 8
source("R/get_change_by_pre.R")
d_x_change <- data.frame(
  pre = pre_values,
  label = pre_labels_x
) |>
  pmap(\(pre, label) get_change_by_pre(
    pre,
    label,
    fit3,
    hours = seq(0, 3, length.out = 100)
  )) |>
  bind_rows()

@@

### Manuscript figures
Create figures for the manuscript.
```{r}
#| label: fig-ms-interaction
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: Figure included in manuscript

scale_x_hours2 <- list(
  scale_x_continuous(
  breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3), 
  labels = c(0, "", "30m", "", "1h", "2h", "3h")
  )
)
p_tmp <- (
  (
    p_pre_x + 
      facet_wrap(~pre, ncol = 4) +
      scale_x_hours2 +
      labs(y = "Mood") +
      theme(axis.title.x = element_blank())
  ) / 
    (
      p_pre_x_change +
        scale_x_hours2 +
        labs(y = "Change", x = "Duration") +
        theme()
    ) 
) +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "none") 
p_tmp
```

```{r}
#| label: ggsave-fig-ms-interaction
#| echo: false

ggsave(
  "figures/fig-pre-x-change.png", 
  p_tmp, 
  width = 7, 
  height = 3.5
  )
```

@@

```{r}
#| label: fig-ms-no-interaction
#| fig-height: 3
#| fig-width: 6
#| fig-cap: Figure shown in manuscript

p_tmp <- (
  p_curve + labs(x = NULL, y = "Mood")
  ) / (
  p_change + labs(x = "Duration", y = "Change")
  ) + plot_annotation(tag_levels = "A") 
p_tmp
```

```{r}
#| label: ggsave-fig-ms-no-interaction
#| echo: false

ggsave(
  "figures/fig-change.png", 
  p_tmp, 
  width = 6, 
  height = 3
  )
```

@@

get_trend_by_model <- function(fit, model_label) {
  emm <- emmeans(
    fit,
    ~ hours + pre,
    at = list(
      hours = seq(0, 3, length.out = 100),
      pre = pre_values
    ),
    lmer.df = "asymp"
  )
  as.data.frame(emm) |>
    mutate(
      pre = factor(
        pre,
        labels = pre_labels_x,
      ),
      model = model_label
    )
}
tmp_p <- map2(
  tmp_models,
  names(tmp_models),
  \(fit, lab) get_trend_by_model(fit, lab)
) |>
  bind_rows()
p_pre_x_sens <- tmp_p |>
  ggplot(
    aes(
      hours,
      emmean,
      group = interaction(model, pre),
      color = model,
      linetype = pre
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      fill = model,
      color = NULL
    ),
    alpha = 0.10,
    show.legend = FALSE
  ) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  scale_x_hours
p_pre_x_sens
```

### 2023-10-06 wrangle.qmd
Cross reference sections (#11)
WIP

---
title: Wrangle
---

@@

library(patchwork)
library(tidyverse)
source("R/functions.R")
```

@@

dat <- read_csv(
  "data/study_prompt_answered.csv",
  col_select = c(
    pid,
    time = Time_utc,
    duration = CurrentSessionLength,
    prompt = LastStudyPromptType,
    mood = response
  )
)
dat <- dat |> 
  mutate(
    pid = factor(pid),
    # Mood to 0-100 scale
    mood = mood / 10, 
    # hours indicates session duration in hours
    hours = duration / 60,
    .keep = "unused"
  )
```

@@

```{r}
#| label: session-indicator-create
dat <- dat |> 
  # filter(hours <= 5) |> 
  arrange(pid, time) |> 
  mutate(
    new_session = hours < lag(hours, default = 999),
    session = cumsum(new_session),
    .by = pid
  ) |> 
  select(-c(new_session)) |> 
  mutate(ps = paste(pid, session, sep = "_"))

@@

```{r}
# What proportion of sessions have a login mood prompt?
dat |> 
  mutate(
    has_login_prompt = any(hours == 0 & prompt == "Wellbeing"),
    .by = ps
  ) |> 
  distinct(ps, has_login_prompt) |> 
  summarise(mean(has_login_prompt))
```

@@

@@

```{r}
dir.create("data", FALSE)
saveRDS(dat, "data/data.rds")
```

>>>>

# Wrangle {#sec-wrangle}

@@

library(patchwork)
library(tidyverse)
source("R/functions.R")
source("R/common.R")
```

@@

dat <- read_csv(
  "data/study_prompt_answered.csv",
  col_select = c(
    pid,
    time = Time_utc,
    duration = CurrentSessionLength,
    prompt = LastStudyPromptType,
    mood = response
  )
)
dat <- dat |> 
  mutate(
    pid = factor(pid),
    # Mood to 0-1 scale
    mood = mood / 1000, 
    # hours indicates session duration in hours
    hours = duration / 60,
    .keep = "unused"
  )
```

@@

```{r}
#| label: session-indicator-create

dat <- dat |> 
  arrange(pid, time) |> 
  mutate(
    new_session = hours < lag(hours, default = 999),
    session = cumsum(new_session),
    .by = pid
  ) |> 
  select(-c(new_session)) |> 
  mutate(ps = paste(pid, session, sep = "_"))

@@

```{r}
#| label: check-login-prompt-rate

# What proportion of sessions have a login mood prompt?
dat |> 
  mutate(
    has_login_prompt = any(hours == 0 & prompt == "Wellbeing"),
    .by = ps
  ) |> 
  distinct(ps, has_login_prompt) |> 
  summarise(mean(has_login_prompt))
```

@@

Other stuff

```{r}
dat <- dat |>
  mutate(
    post = factor(
      row_number() > 1,
      levels = c(FALSE, TRUE),
      labels = c("0", "1")
    ),
    .by = c(pid, session)
  ) |>
  # Censoring indicator
  mutate(
    cl = case_when(
      mood == 0 ~ "left",
      mood == 1 ~ "right",
      TRUE ~ "none"
    )
  )

contrasts(dat$post) <- c(-0.5, 0.5)
```

@@

```{r}
#| label: data-save

dir.create("data", FALSE)
saveRDS(dat, "data/data.rds")
```
